Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-17 numerical-chemistry.tex
Adding note about the variable H2 gamma.

The primordial non-equilibrium chemistry routines used in \enzo\ were first
described by \citet{abel97} and \citet{anninos97}, but have
since been extended with updated reaction rates and the inclusion of
deuterium species \citep{2009PhDT.........5T}.  These routines follow the
non-equilibrium chemistry of a gas of primordial composition with 12 total
species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp, and HD.  \enzo\
also computes the radiative heating and cooling of the gas from atomic and
molecular line excitation, recombination, collisional excitation, free-free
transitions, Compton scattering of the cosmic microwave background, as well as
several models for a metagalactic UV background that heat the gas via
photoionization and photodissociation; see Section~\ref{sec.num.cooling} for
more details.  The chemical and thermal states of the gas can be updated either
at the same hydrodynamical timestep (i.e., de-coupled and operator split) or
through the same subcycling system (i.e., a coupled chemical and thermal
system).  The default behavior of \enzo\ is to couple these two systems at
subcycles of the hydrodynamic timestep; this results in updates to both the
chemical and thermal states of the gas (which also inform the temperature, the
reaction rate coefficients and the cooling functions of the gas) on timescales
that are faster than those of the gas dynamics.

@@

\begin{table}
\begin{center}
{\bfseries Radiative Processes}\\[1ex]
\begin{tabular}{llllllll}
(33) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
(34) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
(35) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
(36) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
(37) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(38) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
(39) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
(40) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
(41) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & \\
(42) & D & + & $\gamma$ & $\rightarrow$ & D$^+$ & + & e$^-$ \\
\end{tabular}
\caption[]{Radiative processes solved in the \enzo\ nonequilibrium
primordial chemistry routines. \red{Missing rates for HD destruction/ionization?}}
\label{table.radiative}
\end{center}
\end{table}

>>>>

The primordial non-equilibrium chemistry routines used in \enzo\ were first
described by \citet{abel97} and \citet{anninos97}, but have
since been extended with updated reaction rates and the inclusion of
deuterium species \citep{2009PhDT.........5T}.  These routines follow the
non-equilibrium chemistry of a gas of primordial composition with 12 total
species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp, and HD.  \enzo\
also computes the radiative heating and cooling of the gas from atomic and
molecular line excitation, recombination, collisional excitation, free-free
transitions, Compton scattering of the cosmic microwave background, as well as
several models for a metagalactic UV background that heat the gas via
photoionization and photodissociation; see Section~\ref{sec.num.cooling} for
more details.  The chemical and thermal states of the gas can be updated either
at the same hydrodynamical timestep (i.e., de-coupled and operator split) or
through the same subcycling system (i.e., a coupled chemical and thermal
system).  The default behavior of \enzo\ is to couple these two systems at
subcycles of the hydrodynamic timestep; this results in updates to both the
chemical and thermal states of the gas (which also inform the temperature, the
reaction rate coefficients and the cooling functions of the gas) on timescales
that are faster than those of the gas dynamics.  The gamma used by \enzo\ to
compute the temperature of the gas from the energy and density characteristics
utilizes a variable gamma that includes effects of the rotational state of
molecular hydrogen, enabling it to vary from $5/3$ (fully-atomic) to $7/5$
(fully molecular).  This further coupling of the chemical and thermal states of
the gas underscores the need for coupled chemistry and radiative cooling
solutions.

@@

\begin{table}
\begin{center}
{\bfseries Radiative Processes}\\[1ex]
\begin{tabular}{llllllll}
(33) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
(34) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
(35) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
(36) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
(37) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(38) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
(39) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
(40) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
(41) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & \\
(42) & D & + & $\gamma$ & $\rightarrow$ & D$^+$ & + & e$^-$ \\
\end{tabular}
\caption[]{Radiative processes solved in the \enzo\ nonequilibrium
primordial chemistry routines.}
%\red{Missing rates for HD destruction/ionization?}}
\label{table.radiative}
\end{center}
\end{table}
